Tensor denoising and completion based on ordinal observations

### 2020-03-05 paper/arxiv/correction.tex
Hession bound

Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1},\ldots,\mz_{d_1,\ldots,d_K})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$. It follows from property (a) that
\[
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega.
\]
Furthermore, properties (b) and (c) (or similar arguments as in page 29, arXiv preprint) imply that
\begin{equation}
\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\KeepStyleUnderBrace{\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}}_{\text{Bernoulli r.v.\ with probability bounded by $A_{\alpha,\beta,\Delta}$}}.
\end{equation}
Therefore,
\begin{equation}\label{eq:sum}
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}.
\end{equation}
Based on central limit theorem, as the tensor dimension goes to infinity, 
\begin{equation}\label{eq:clt}
\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \rightarrow\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathbb{P}(y_\omega=\ell \text{ or }\ell+1)\geq c_2A_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2
\end{equation}
holds with very high probability. 
By~\eqref{eq:sum} and~\eqref{eq:clt}, we have
\begin{equation}\label{eq:boundH}
\mz^T\mH\mz \leq-c_2A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
\end{equation}
and therefore~\eqref{eq:HessionZ} is proved. Plugging~\eqref{eq:scoreZ} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
\[
\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K-1}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
\]
The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}. 
\end{proof}

\bibliography{tensor_wang}
\bibliographystyle{apalike}
\end{document}

>>>>

Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1},\ldots,\mz_{d_1,\ldots,d_K})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$. It follows from property (a) that
\[
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega.
\]
Furthermore, properties (b) and (c) (or similar arguments as in page 29, arXiv preprint) imply that
\begin{equation}
\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\KeepStyleUnderBrace{\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}}_{\text{Bernoulli r.v.\ with probability bounded by $A_{\alpha,\beta,\Delta}$}}.
\end{equation}
Therefore,
\begin{equation}\label{eq:sum}
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}.
\end{equation}

{\color{red}
Based on central limit theorem (and concentration properties of Bernoulli r.v.'s), as the tensor dimension goes to infinity, 
\begin{equation}\label{eq:clt}
\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \rightarrow\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathbb{P}(y_\omega=\ell \text{ or }\ell+1)\geq c_2A_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2
\end{equation}
holds with very high probability. 
}

By~\eqref{eq:sum} and~\eqref{eq:clt}, we have
\begin{equation}\label{eq:boundH}
\mz^T\mH\mz \leq-c_2A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
\end{equation}
and therefore~\eqref{eq:HessionZ} is proved. Plugging~\eqref{eq:scoreZ} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
\[
\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K-1}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
\]
The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}.

{ \color{red}Correction of~\eqref{eq:clt}. Define the subspace:
\[
\tS=\{\Vec(\tZ): \tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,\ (\Theta,\mb)\in(\tP,\tB)\}.
\]
We show that Hession is definite negative restricted in the subspace $\tS$. Specifically, for any vector $\mz=\entry{z_{\omega,\ell}}\in\tS$,
\begin{align}
\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}&=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
&=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}+\sum_{\omega,\ell}b^2_\ell\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
&\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
&\geq \sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2
\end{align}

On the other hand,
\[
\FnormSize{}{\mz}^2=\sum_{\omega,\ell}z^2_{\omega,\ell}=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2=(L-1)\sum_{\omega}\theta^2_\omega+\left(\prod_k d_k\right)\sum_\ell b^2_\ell.
\]

Therefore, there exists a positive constant $c_1>0$ (? perhaps depending on $\alpha, \beta,\Delta$ etc...needs some caculation..) such that
\[
\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \over \FnormSize{}{\mz}^2}\geq c_1 {\min_{\ell}(n_\ell+n_{\ell+1})\over (L-1)\prod_k d_k}.
\]
The conclusion follows by noting that the ratio ${\min_{\ell}(n_\ell+n_{\ell+1})\over \prod_k d_k} \geq c'A_{\alpha,\beta,\Delta}$ in high probability. 
}

\end{enumerate}1
\end{proof}
\bibliography{tensor_wang}
\bibliographystyle{apalike}
\end{document}
